MA_Results

### 2023-10-16 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
close to submission update - except tables

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., in a format that allows computers to perform calculations). The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. However, the ease of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.
<!-- through execution of a computer program that applies the computational model to selected input data). -->
<!-- the execution of a computer program that applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations. -->
<!-- the predictions reported in health economic studies are now almost always -->

The set of digital files implementing a health economic model's structure, data and algorithms as software can be called a computational health economic model (CHEM). CHEMs can be developed using specialised commercial modelling software or in an open-source programming language such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial modelling software is that users often require no or limited programming skills to develop and apply CHEMs. However, open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what software development approach to adopt needs to balance both modelling team value judgments about priority CHEM features and the skills, resource and time constraints within which a CHEM project needs to be delivered (see Table \@ref(tab:proscons)).

@@

Within the health economics discipline, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. The main barriers to open-source CHEMs identified in a 2020 survey of health economists concern challenges relating to updating models, legal considerations, transferring model data, the platforms or languages in which models are developed, the level of public access required, lack of interest from decision-makers, model generalisability and confidentiality and security considerations [@Pouwels2022]. 
<!-- A major consideration in open-source software development, is the selection of an appropriate software framework.  -->
<!-- on considerations such as model comprehensiveness, generalisability, reproducibility and transferability -->

Some of these barriers could potentially be mitigated through the development of appropriate software frameworks. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

@@

A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. A software framework that help CHEM developers fulfill a broader set of ethical responsibilities would require a number of conceptual, technical and behavioural challenges to be overcome. Firstly, the ethical responsibilities of developers of computational models designed to inform governmental decision-making are often poorly understood <!-- and inadequately fulfilled  -->[@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational]. Software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Furthermore, software frameworks can be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development].
<!-- a software framework is unlikely to be widely adopted unless the benefits to developers outweigh the costs associated with using it. -->
<!-- However, guidance on developing, selecting and using software frameworks for ethical CHEM implementations is scarce. -->

>>>>

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., the format that allows computers to perform calculations). The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. However, the ease of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.
<!-- through execution of a computer program that applies the computational model to selected input data). -->
<!-- the execution of a computer program that applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations. -->
<!-- the predictions reported in health economic studies are now almost always -->

The set of digital files implementing a health economic model's structure, data and algorithms as software can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling software or in an open-source programming language such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial modelling software is that users often require no or limited programming skills to develop and apply CHEMs. However, open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what type of software development platform to use needs to balance ethical and pragmatic considerations  (see Table \@ref(tab:proscons)).

<!-- modelling team value judgments about priority CHEM features and the skills, resource and time constraints within which a CHEM project needs to be delivered -->

@@

Open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. Additional resources, skills and tools may be required to overcome a number of barriers to greater health economist adoption of open-source development. CHEM software development platforms were identified in a 2020 survey of health economists as a barrier to open-source model implementations  [@Pouwels2022]. Other reported barriers relate to difficulties in generalising models, updating models and transferring model data, concerns about the level of public access required and lack of interest from decision-makers, and legal, confidentiality and security considerations [@Pouwels2022].

<!-- A major consideration in open-source software development, is the selection of an appropriate software framework.  -->
<!-- on considerations such as model comprehensiveness, generalisability, reproducibility and transferability -->
<!-- Although there appears to be in principle support from many health economists for greater use of open source health economic models (OSHEMs)  [@Pouwels2022], actual implementations are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption of OSHEMs include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020],  -->

Better software frameworks could potentially improve open-source CHEM development. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

@@

A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, to gain widespread adoption, a software framework for CHEM developers may need to provide tools to help meet a broader range of ethical practice requirements. There are multiple challenges to overcome. Firstly, the ethical responsibilities of public health computational modellers remain poorly understood <!-- and inadequately fulfilled  [@pliakos2021ethics; @thompson2022escape;]-->[@10.3389/fpubh.2017.00068] and the computational modelling landscape is changing fast [@calder2018computational. Secondly, software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Thirdly, software frameworks can be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development].
<!-- a software framework is unlikely to be widely adopted unless the benefits to developers outweigh the costs associated with using it. -->
<!-- However, guidance on developing, selecting and using software frameworks for ethical CHEM implementations is scarce. -->

### 2023-10-16 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
close to submission update - except tables

# Assessing ethical CHEM implementations
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

# Software framework
```{r child="../Child_RMDs/MS_Imp_Toolkit.Rmd", echo=FALSE}
```

# Application
```{r child="../Child_RMDs/MS_Imp_Example.Rmd", echo=FALSE}
```

>>>>

# Ethical responsibilities and assessment criteria
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

# Prototype software framework
```{r child="../Child_RMDs/MS_Imp_Toolkit.Rmd", echo=FALSE}
```

# Application in youth mental health
```{r child="../Child_RMDs/MS_Imp_Example.Rmd", echo=FALSE}
```
